Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-06-26 optim.qmd
SBO acronym added

The paper is structured as follows. @sec-background introduces the background of PP guided tour and reviews existing optimisers and index functions in the literature. @sec-theory details JSO and introduces metrics that measure different properties of PP indexes. @sec-vis outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's performance. @sec-app presents two sets of simulation experiments, comparing JSO with another optimiser, search better, and studies the impact of different PP index properties on the optimisation
performance. @sec-conclusion summarises the work and provides suggestions for future directions.

@@

The performance of JSO is investigated in two folds: 1) compare it with
a commonly used optimiser: the search-better optimiser [@RJ-2021-105;
@laa_using_2020], and 2) examine its success rate under different
hyper-parameters (number of jellyfishes and maximum number of tries).
The projection pursuit problem used is finding the pipe shape using the
holes index, investigated by @laa_using_2020.

```{r}
#| fig.height = 7.5,
#| fig.width = 10,
#| label = fig-proj,
#| fig.align = "center",
#| fig.cap = "Projections found by the jellyfish and search better optimisers at each 10th quantile across 50 simulations. The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces. The JSO uses 100 jellyfishes and a maximum number of tries of 100. The search better optimiser uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates. In the 6-D data space, JSO always finds a clear pipe shape while the search better optimiser also finds the pipe shape but with a wide rim. At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the search better optimiser."

@@

(p1 + ggtitle("The jellyfish search optimiser"))/(p2 + ggtitle("The search better optimiser"))
```

@@

Fifty simulations are conducted with both JSO and the search better optimal, in
four dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes and
100 maximum number of tries, while the search-better optimiser allows a
maximum of 1000 samples at each iteration before the algorithm
terminates. @fig-proj presents the final projections found by the two
optimisers, broken down by 10th quantile, faceted by the data dimension.
In the 6-dimensional data scenario, JSO consistently identifies a clear
pipe shape. The search better optimiser also finds the pipe shape but
with a wide rim, suggesting a further polish search may be required.
With increasing dimensions, JSO may not always identify the pipe shape
due to random sampling, but it still presents a pipe shape in over 50%
of cases. When compared to the search better optimiser, JSO reaches
higher index values and clearer pipe shapes across all quantiles.

>>>>

The paper is structured as follows. @sec-background introduces the background of PP guided tour and reviews existing optimisers and index functions in the literature. @sec-theory details JSO and introduces metrics that measure different properties of PP indexes. @sec-vis outlines the data structures, numerical summaries, and visualisations used in the simulation studies conducted in @sec-app to assess JSO's performance. @sec-app presents two sets of simulation experiments, comparing JSO permformance improvements relative to an existing optimiser, and studies the impact of different PP index properties on the optimisation performance. @sec-conclusion summarises the work and provides suggestions for future directions.

@@

The performance of JSO is investigated in two folds: 1) compare it with a commonly used optimiser: the search-better optimiser (SBO) [@RJ-2021-105;
@laa_using_2020], and 2) examine its success rate under different hyper-parameters (number of jellyfishes and maximum number of tries). The projection pursuit problem used is finding the pipe shape using the holes index, investigated by @laa_using_2020.

```{r}
#| fig.height = 7.5,
#| fig.width = 10,
#| label = fig-proj,
#| fig.align = "center",
#| fig.cap = "Projections found by the JSO and SBO at each 10th quantile across 50 simulations. The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces. The JSO uses 100 jellyfishes and a maximum number of tries of 100. The SBO uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates. In the 6-D data space, JSO always finds a clear pipe shape while the SBO also finds the pipe shape but with a wide rim. At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the SBO."

@@

(p1 + ggtitle("a. JSO"))/(p2 + ggtitle("b. SBO"))
```

@@

Fifty simulations are conducted with both JSO and the SBO, in four dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes and 100 maximum number of tries, while the SBO allows a maximum of 1000 samples at each iteration before the algorithm terminates. @fig-proj presents the final projections found by the two optimisers, broken down by 10th quantile, faceted by the data dimension. In the 6-dimensional data scenario, JSO consistently identifies a clear pipe shape. The SBO also finds the pipe shape but with a wide rim, suggesting a further polish search may be required. With increasing dimensions, JSO may not always identify the pipe shape due to random sampling, but it still presents a pipe shape in over 50%
of cases. When compared to the SBO JSO reaches higher index values and clearer pipe shapes across all quantiles.

### 2024-06-26 optim.tex
SBO acronym added

The paper is structured as follows. Section~\ref{sec-background}
introduces the background of PP guided tour and reviews existing
optimisers and index functions in the literature.
Section~\ref{sec-theory} details JSO and introduces metrics that measure
different properties of PP indexes. Section~\ref{sec-vis} outlines the
data structures, numerical summaries, and visualisations used in the
simulation studies conducted in Section~\ref{sec-app} to assess JSO's
performance. Section~\ref{sec-app} presents two sets of simulation
experiments, comparing JSO with another optimiser, search better, and
studies the impact of different PP index properties on the optimisation
performance. Section~\ref{sec-conclusion} summarises the work and
provides suggestions for future directions.

@@

\begin{tcolorbox}[enhanced jigsaw, opacityback=0, coltitle=black, colframe=quarto-callout-note-color-frame, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, breakable, bottomtitle=1mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, toptitle=1mm, opacitybacktitle=0.6, bottomrule=.15mm, left=2mm, colback=white, titlerule=0mm, toprule=.15mm]

@@

The performance of JSO is investigated in two folds: 1) compare it with
a commonly used optimiser: the search-better optimiser
\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
under different hyper-parameters (number of jellyfishes and maximum
number of tries). The projection pursuit problem used is finding the
pipe shape using the holes index, investigated by
\citet{laa_using_2020}.

@@

\caption{\label{fig-proj}Projections found by the jellyfish and search
better optimisers at each 10th quantile across 50 simulations. The
projection pursuit problem is to find the pipe shape using the holes
index in the 6, 8, 10, and 12-dimensional spaces. The JSO uses 100
jellyfishes and a maximum number of tries of 100. The search better
optimiser uses a maximum of 1000 tries in each step of random sampling
step before the algorithm terminates. In the 6-D data space, JSO always
finds a clear pipe shape while the search better optimiser also finds
the pipe shape but with a wide rim. At higher data dimensions, JSO finds
a higher index value and a clearer pipe shape across all the quantiles
than the search better optimiser.}

@@

Fifty simulations are conducted with both JSO and the search better
optimal, in four dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100
jellyfishes and 100 maximum number of tries, while the search-better
optimiser allows a maximum of 1000 samples at each iteration before the
algorithm terminates. Figure~\ref{fig-proj} presents the final
projections found by the two optimisers, broken down by 10th quantile,
faceted by the data dimension. In the 6-dimensional data scenario, JSO
consistently identifies a clear pipe shape. The search better optimiser
also finds the pipe shape but with a wide rim, suggesting a further
polish search may be required. With increasing dimensions, JSO may not
always identify the pipe shape due to random sampling, but it still
presents a pipe shape in over 50\% of cases. When compared to the search
better optimiser, JSO reaches higher index values and clearer pipe
shapes across all quantiles.

>>>>

The paper is structured as follows. Section~\ref{sec-background}
introduces the background of PP guided tour and reviews existing
optimisers and index functions in the literature.
Section~\ref{sec-theory} details JSO and introduces metrics that measure
different properties of PP indexes. Section~\ref{sec-vis} outlines the
data structures, numerical summaries, and visualisations used in the
simulation studies conducted in Section~\ref{sec-app} to assess JSO's
performance. Section~\ref{sec-app} presents two sets of simulation
experiments, comparing JSO permformance improvements relative to an
existing optimiser, and studies the impact of different PP index
properties on the optimisation performance. Section~\ref{sec-conclusion}
summarises the work and provides suggestions for future directions.

@@

\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, opacitybacktitle=0.6, toprule=.15mm, coltitle=black, colback=white, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, opacityback=0, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, toptitle=1mm, bottomtitle=1mm, titlerule=0mm, left=2mm]

@@

The performance of JSO is investigated in two folds: 1) compare it with
a commonly used optimiser: the search-better optimiser (SBO)
\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
under different hyper-parameters (number of jellyfishes and maximum
number of tries). The projection pursuit problem used is finding the
pipe shape using the holes index, investigated by
\citet{laa_using_2020}.

@@

\caption{\label{fig-proj}Projections found by the JSO and SBO at each
10th quantile across 50 simulations. The projection pursuit problem is
to find the pipe shape using the holes index in the 6, 8, 10, and
12-dimensional spaces. The JSO uses 100 jellyfishes and a maximum number
of tries of 100. The SBO uses a maximum of 1000 tries in each step of
random sampling step before the algorithm terminates. In the 6-D data
space, JSO always finds a clear pipe shape while the SBO also finds the
pipe shape but with a wide rim. At higher data dimensions, JSO finds a
higher index value and a clearer pipe shape across all the quantiles
than the SBO.}

@@

Fifty simulations are conducted with both JSO and the SBO, in four
dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes and 100
maximum number of tries, while the SBO allows a maximum of 1000 samples
at each iteration before the algorithm terminates. Figure~\ref{fig-proj}
presents the final projections found by the two optimisers, broken down
by 10th quantile, faceted by the data dimension. In the 6-dimensional
data scenario, JSO consistently identifies a clear pipe shape. The SBO
also finds the pipe shape but with a wide rim, suggesting a further
polish search may be required. With increasing dimensions, JSO may not
always identify the pipe shape due to random sampling, but it still
presents a pipe shape in over 50\% of cases. When compared to the SBO
JSO reaches higher index values and clearer pipe shapes across all
quantiles.
